New Bayes covariant coupling methods to cope with change of priors in one-vs-one classification

### 2025-08-10 experiment.tex
Fix from Peter's proof

\section{Experiments.}

The choice of a coupling method is an important decision for practitioners. We aim to investigate the following interrelated questions

\begin{itemize}
	\item[a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu-Lin-Weng (see Section \ref{sec:exp1})? 
	\item[b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance ((see Section \ref{sec:exp2} and \ref{sec:exp3}))?
	\item[c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice ((see Section \ref{sec:exp4}))?
\end{itemize}

These questions will be put to test in Sections \ref{sec:exp1} --\ref{sec:exp4}. The methodology of our experiments is described in the next two subsections.

@@

It is natural to consider the benchmark datasets used in previous evaluation of coupling methods \cite{wu2004probability}. The authors considered altogether seven datasets with number of classes ranging from 3 to 26. For each dataset they extracted 40 pairs of training and testing datasets. Half of these were smaller (300 training samples and 500 testing samples) and the other half was larger (800 training samples and 1000 testing samples). In our experiments we use only the larger pairs.

@@

We can infer that except for 'waveform' dataset, majority of binary problems are linearly separable.

@@

We have opted to use LDA as a binary classification tool. Its appeal is that it is applicable even for linearly separable datasets, thus obviating additional crossvalidation step incurred by methods that use regularization (e.g. penalized logistic regression or SVM). It usually provides results close to logistic regression [\cite{james2013introduction}].

@@

For training LDA we used MASS library in R. We implemented coupling methods in C++ using Eigen library. As arguments to the coupling methods we use log-odds, i.e. instead of $r_{ij}$ they expect
$$
\tilde r_{ij} = \log \biggl(\frac{1}{r_{ji}} - 1\biggr).
$$
The resulting matrix $\tilde{\boldsymbol{R}}$ is skew symmetric (with the convention of having zeroes on the diagonal).

@@

Small probabilities generally have very little effect on the implementation of Bayes covariant methods which generally operate using log-odds. However small probabilities do affect both the method of Wu-Lin-Weng as well sa radial method, because then Assumption \ref{ass:1} fails to hold.

@@

The first experiment compares the performance the following canonical coupling methods
\begin{itemize}
\item method of Wu-Lin-Weng, which is widely used in software libraries, and thus can be viewed as an orthodox choice,
\item normal coupling, which is a Bayes covariant method,
\item radial coupling, which is not Bayes covariant,
\item Hinton's oracle.
\end{itemize}

@@

Let us define two parameter family of Bayes covariant coupling methods
\begin{align}
a_1 \boldsymbol{s}_1 \oplus a_2 \boldsymbol{s}_2 \oplus a_3 \boldsymbol{s}_3,\quad\textrm{where } a_1 + a_2 +a_3 = 1. \label{eq:family}
\end{align}
Note that this family includes the normal coupling method \eqref{eq:normal3} as well as the three arboreal methods $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$.

@@

Selecting parameters for methods in family \eqref{eq:family} requires several methodological choices. 
\begin{itemize}
\item To avoid bias in selection of parameters $a_1, a_2, a_3$ we use a subset of testing dataset to find their optimal values. As a result, we will be comparing the accuracy of Wu-Lin-Weng method evaluated with repeated five-fold cross-validation accuracy of a method in family \eqref{eq:family}. We use 20 repetitions.
\item To compare two parameter vectors, we compare their resulting cross-validation accuracy. An alternative approach would be to use  a gradient based optimization for a smooth proxy (e.g. cross-entropy) of empirical accuracy. We opted for the former because the problem is two-dimensional, and we can use grid search instead.
\item Our choice of grid search necessitates a selection of a compact subset of parameter space that will be sampled. We choose to examine points in the simplex defined by  $a_i\geq 0$, for $i=1,2,3$. As Table \ref{tab:step2} shows, this simplex contained in more than half case a Bayes covariant method matching the performance of Wu-Lin-Weng's method.
\end{itemize}

>>>>

\section{Experiments}

The choice of a coupling method is an important decision for practitioners. We aim to investigate the following interrelated questions:

\begin{itemize}
	\item[(a)] Which coupling methods without parameters show good performance across diverse data sets? Is it the orthodox method of Wu-Lin-Weng? An experiment to answer this question is carried out Section \ref{sec:exp1}. 
	\item[(b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric Bayes covariant method which  matches its performance? An experiment related to this question is carried out in  Sections \ref{sec:exp2} and \ref{sec:exp3}.
	\item[(c)] If the answer to b) is negative, is it better to apply the change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice. This question is investigated in Section \ref{sec:exp4}.
\end{itemize}

These questions will be put to test in Sections \ref{sec:exp1}--\ref{sec:exp4}. The methodology of our experiments is described in the next two subsections.

@@

It is natural to consider the benchmark datasets used in previous evaluation of coupling methods \cite{wu2004probability}. The authors considered altogether seven datasets with number of classes ranging from 3 to 26. For each dataset they extracted 40 pairs of training and testing datasets. Half of these were smaller (300 training samples and 500 testing samples) and the other half were larger (800 training samples and 1000 testing samples). In our experiments we use only the larger pairs.

@@

We can infer that except for 'waveform' dataset, majority of binary subproblems are linearly separable.

@@

We have opted to use LDA as a binary classification tool. Its appeal is that it is applicable even for linearly separable datasets, thus obviating additional crossvalidation step incurred by methods that use regularization (e.g. penalized logistic regression or SVM). It usually provides results close to logistic regression \cite{james2013introduction}.

@@

For training LDA we used MASS library in R. We implemented coupling methods in C++ using Eigen library. As arguments to the coupling methods we use log-odds:
$$
\tilde r_{ij} = \log \biggl(\frac{1}{r_{ji}} - 1\biggr).
$$
The resulting matrix $\tilde{\boldsymbol{R}}$ is skew symmetric (with the convention of having zeroes on the diagonal).

@@

Small probabilities generally have very little effect on the implementation of Bayes covariant methods which generally operate using log-odds. However small probabilities do affect both the method of Wu-Lin-Weng as well as radial method, because then Assumption \ref{ass:2} fails to hold.

@@

The first experiment compares the performance of the following canonical coupling methods:
\begin{itemize}
\item method of Wu-Lin-Weng, which is widely used in software libraries, and thus can be viewed as an orthodox choice,
\item normal coupling, which is Bayes covariant,
\item radial coupling, which is not Bayes covariant,
\item Hinton's oracle.
\end{itemize}

@@

Let us define a two-parameter family of Bayes covariant coupling methods
\begin{align}
a_1 \boldsymbol{s}_1 \oplus a_2 \boldsymbol{s}_2 \oplus a_3 \boldsymbol{s}_3,\quad\textrm{where } a_1 + a_2 +a_3 = 1. \label{eq:family}
\end{align}
Note that this family includes the normal coupling method \eqref{eq:normal3} as well as the three arboreal methods $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$.

@@

Selecting parameters for methods in family \eqref{eq:family} requires several methodological choices. 
\begin{itemize}
\item To avoid bias in selection of parameters $a_1, a_2, a_3$ we use a subset of testing dataset to find their optimal values. As a result, we will be comparing the accuracy of Wu-Lin-Weng method evaluated with repeated five-fold cross-validation accuracy of a method in family \eqref{eq:family}. We use 20 repetitions.
\item To compare two parameter vectors, we compare their resulting cross-validation accuracy. An alternative approach would be to use  a gradient based optimization for a smooth proxy (e.g. cross-entropy) of empirical accuracy. We opted for the former because the problem is two-dimensional, and we can use grid search instead.
\item Our choice of grid search necessitates a selection of a compact subset of parameter space that will be sampled. We choose to examine points in the simplex defined by  $a_i\geq 0$, for $i=1,2,3$. As Table \ref{tab:step2} shows, this simplex contained in more than half cases a Bayes covariant method matching the performance of Wu-Lin-Weng's method.
\end{itemize}
